---
title: "Data 607 Tidyverse Create Assignment"
author: "Wilson Chau"
date: "2022-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
I took this dataset from 
Tidyverse packages consists of many different uses. These are the few I looked up:
ggplot
dplyr
tidyr
readr
purrr
tibble
stringr
forcats

I wanted to focus on dplyr because of the package's abiltiy to dtransform and summarize tabular data with row. I also wanted to focus more on having descriptive data with some numerical data to play around with.

One set of data I took was the Thanksgiving choice/poll for 2015. 
https://fivethirtyeight.com/features/heres-what-your-part-of-america-eats-on-thanksgiving/
Let's get started on using DPLYR functions. 

```{r}
thanksgiving = read.csv('https://raw.githubusercontent.com/Wilchau/Data_607_tidyverse_create_Thanksgiving/main/thanksgiving-2015-poll-data.csv')
```

Important dplyr verbs to remember:
select()	select columns
filter()	filter rows
arrange()	re-order or arrange rows
mutate()	create new columns
summarise()	summarise values
group_by()	allows for group operations in the “split-apply-combine” concept

I focus on the columns that talks about celebrating Thanksgiving


##Question 1: Can you select and view the following columns: Age group, Household income, and US region? 
Answer: use select()
```{r}
thanksgiving_info <- select(thanksgiving, Age, How.much.total.combined.money.did.all.members.of.your.HOUSEHOLD.earn.last.year., US.Region )
head(thanksgiving_info)
```
#Question 2: Can you find the number of particpants over the age of 30?
A:  Use filter() follow by operator > 
```{r}
thanksgiving_info %>% 
  filter(Age > 30)
```
## Question 3: Can you sort the Age 60+ with highest to lowest income
```{r}
thanksgiving_info %>% 
  filter(Age > 60) %>%
  arrange(((How.much.total.combined.money.did.all.members.of.your.HOUSEHOLD.earn.last.year.)))
```

## Question 4: What is the most celebrated region?
Answer: Use group_by()
```{r}
thanksgiving %>% 
  group_by(Do.you.celebrate.Thanksgiving., US.Region) %>%
  summarize(count=n())
```
